---
title: "PS270 Final Project"
author: "Lillie Claire Sunby"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Project thoughts

I am interested in exploring data related to...
how democracies may impact whether countries have gender quotas. I could look at different kinds of democracies across countries and compare the gender quotas in the countries.

## Final Project Milestone 2

My research question has changed as a result of finding a different data set. My research question is: Does an increase in the adoption of gender quotas increase women’s representation in national legislatures? I will examine whether countries with gender quotas have a higher percentage of women in their lower houses of parliament. Gender quotas are put in place to help women overcome barriers to becoming candidates, and they require a minimum level of female representation. My independent variable is effective.quota, which is coded as 1 if it is present in the country and 0 if it is not. My dependent variable is women.rep, which is the percentage of women in the lower house of parliament. I hypothesize there will be a positive relationship between effective quotas and women’s representation in parliament. If the data shows a positive relationship (in countries with effective quotas, there are higher percentages of female representation), and a statistically significant coefficient for effective.quota, this would support my hypothesis. If the coefficient is not positive or significant, this would suggest that these gender quotas do not increase women’s representation.

```{r}
library(readr)
library(tidyverse)

QAROT_DATA <- read_csv(
  "~/Desktop/PS 270/QAROT_DATA.csv",
  skip = 1
)
head(QAROT_DATA)
```

```{r}
clean_data <- QAROT_DATA |>
  select(country, year, effective.quota, women.rep) |>
  mutate(
    year = as.numeric(year),
    effective.quota = as.numeric(effective.quota),
    women.rep = as.numeric(women.rep)
  ) |>
  drop_na(effective.quota, women.rep)

clean_data
```

```{r}
ggplot(clean_data, aes(x = factor(effective.quota), y = women.rep)) +
  geom_boxplot(fill = "lightblue") +
  labs(
    title = "Effect of Gender Quotas on Women's Representation in Parliament",
    x = "Effective Gender Quota (0 = No, 1 = Yes)",
    y = "Women in Parliament(%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

## Analysis

The boxplot shows women's representation in parliament in countries with an effective gender quota, and without an effective gender quota. As seen above, countries that had effective gender quotas showed higher median levels of women’s representation in comparison to countries without quotas.The distribution for countries with effective quotas is shifted upward. This shows that these quotas are associated with increased representation in parliament. Overall, the boxplot supports my initial hypothesis that gender quotas are associated with higher levels of women’s representation in parliament.

##Results
```{r}
library(tidyverse)

summary_table <- clean_data |>
  group_by(effective.quota) |>
  summarize(
    mean_representation = mean(women.rep),
    median_representation = median(women.rep),
    sd_representation = sd(women.rep),
    n = n(),
    .groups = "drop")

knitr::kable(summary_table, digits = 2)
```
